\documentclass[11pt]{article}
\setlength{\oddsidemargin}{0in}
\setlength{\topmargin}{0.0in}
\setlength{\textwidth}{6.7in}
\setlength{\textheight}{8.5in}
%
%\input mydefs.tex
\def\vev#1{\left\langle #1\right\rangle}
\def\hb{\hfill\break}
%
\begin{document}
%
\centerline{\large\bf Physics 2c \hfill Assignment IV \hfill  4/23/15}

\medskip
\begin{list}{}{\leftmargin 2.4cm \labelsep .2cm \labelwidth 2.2cm}
\item[{\bf Reading:}  \hfill ] This week: K\&K, Chapter 4. \hb
(including the Greenhouse Supplement at the end of the chapter) \hb
                               Next week: K\&K, Chapter 5.
\item[{\bf Problems:} \hfill ] Due in Ph2 IN box, 4\,PM  Thursday, 4/30/15.
\end{list}

\begin{description}

\item[{\bf IV.1} ] {\bf Thermodynamics of the Sun.} K\&K 4.2, 4.3, 4.4.

\item[{\bf IV.2} ] {\bf Surface Temperature of the Earth.} K\&K 4.5. \hb
(The actual surface temperature is some 30\,K warmer due to the Greenhouse effect).

\item[{\bf IV.3} ] {\bf Thermal Radiation Pressure.} K\&K 4.6.

\item[{\bf IV.4} ] {\bf Heat shields.} K\&K 4.8.

\item[{\bf IV.5} ] {\bf Isentropic cosmological expansion.} K\&K 4.18.

\end{description}

\end{document}

\item[{\bf IV.4} ] {\bf Heat capacity of photons and phonons.} K\&K 4.12.

\item[{\bf IV.1} ] {\bf Thermodynamic potentials and work.}
A substance has the following properties:
\begin{itemize}
\item[ ] \vskip -.2cm
\begin{itemize}
\item[{\it i})] 
At a particular constant temperature $\tau_0$,
the mechanical work it does in expanding from $V_0$ to $V$ is
$$W = \int^{V}_{V_0} p\,dV = N\tau_0 \log\left(\frac{V}{V_0}\right) . $$
\item[{\it ii})] 
The entropy at any temperature $\tau$ and volume $V$ is 
$\sigma = N \frac{V_0}{V}\left(\frac{\tau}{\tau_0}\right)^\alpha$,
where $\tau_0$ and $V_0$ are the constants of {\it i})
and $\alpha$ is another constant.
\end{itemize}
\item[{\bf a)}]
Calculate the Helmholtz free energy $F(V,\tau)$ relative to its value
at $F(V_0,\tau_0)$; that is, find $dF = F(V,\tau) - F(V_0,\tau_0)$.
[ Hint: The difference in $F$ between any two points
in the $(V,\tau)$ plane is the line integral of the partial derivatives
of $F$ along any convenient path connecting the two points.
As a proper thermodynamic potential, the difference is independent
of the path taken.]
\item[{\bf b)}]
What is the equation of state?
\item[{\bf c)}]
Find $U(V,\tau)$.
\item[{\bf d)}]
Find the work done in an isothermal expansion from $V_0$ to $V$
at an arbitrary temperature $\tau$.
\end{itemize}

